---
title: "Senescence Significance Testing"
author: ""
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
output: ioslides_presentation
---

<style type="text/css">
@import url('https://fonts.googleapis.com/css2?family=Montserrat:wght@300;400;600&family=Source+Sans+Pro:wght@300;400;600&display=swap');

body, p, li {
  font-family: 'Source Sans Pro', sans-serif;
  color: #333333;
  font-size: 20px;
  line-height: 1.5;
}

h1, h2, h3, h4, h5 {
  font-family: 'Montserrat', sans-serif;
  font-weight: 600;
}

slides > slide.title-slide {
  background: linear-gradient(to bottom right, #8C1D40, #4B2E83);
}

slides > slide.title-slide hgroup h1 {
  color: #FFFFFF;
  font-size: 52px;
  text-shadow: 1px 1px 3px rgba(0,0,0,0.3);
  margin-bottom: 30px;
}

slides > slide.title-slide hgroup h2 {
  color: #FFFFFF;
  font-size: 32px;
}

h2 {
  color: #8C1D40;
  border-bottom: 2px solid #8C1D40;
  padding-bottom: 5px;
}

slides > slide {
  background-color: #FFFFFF !important; /* Force white background */
  border-top: 5px solid #8C1D40;
}

slides > slide:before {
  background: none !important;
}

slides > slide p, slides > slide li {
  color: #333333 !important; /* Darker text for better readability */
}

em, i {
  font-style: italic;
  color: #333333;
}

pre {
  background-color: #f8f8f8;
  border-left: 5px solid #8C1D40;
  font-size: 16px;
}

table {
  width: 100%;
  border-collapse: collapse;
  margin-bottom: 20px;
}

table th {
  background-color: #8C1D40;
  color: white;
  padding: 10px;
}

table td {
  padding: 8px;
  border: 1px solid #ddd;
}

table tr:nth-child(even) {
  background-color: #f2f2f2;
}

ul, ol {
  margin-left: 20px;
}

li {
  margin-bottom: 10px;
}

strong {
  color: #8C1D40;
  font-weight: 600;
}

img, .plotly, svg {
  border-radius: 5px;
  box-shadow: 0 4px 6px rgba(0,0,0,0.1);
  margin: 10px 0;
}

slides > slide:not(.nobackground):after {
  content: '';
  background: none;
}

slides > slide::after {
  content: 'Senescence Significance Testing - ' attr(data-slide-num) ' / ' attr(data-total-slides);
  font-size: 12px;
  position: absolute;
  bottom: 10px;
  right: 20px;
  color: #666;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(plotly)
library(reshape2)
library(stats)
library(knitr)
library(broom)
library(corrplot)
```

# What is Cellular Senescence?

Cellular senescence is a state in which cells permanently stop dividing in response to stress or damage. The term "senescence" comes from the Latin <em>senescere</em>, meaning "to grow old."

```{r data, include=FALSE}
cellage_data <- read.csv("data/genage_human.csv")
summary(cellage_data[, 1:5])
```
